LLM-Driven CIDOC-CRM Knowledge Graph Extraction from Museum Archives

### 2025-05-02 paper/paper.tex
paper: revise IV experiment and VI conclusion

\section{Introduction}

@@

\section{Related Work}

@@

\section{Methods}

@@

\section{Experiment}

We perform ontology-free and ontology-guided graph extraction on open-weight LLMs of various sizes, producing two graphs for each model. Specifically, we conducted experiments using Llama 3 \cite{grattafiori2024llama3herdmodels} (3B and 8B) and Qwen 2.5 \cite{qwen2025qwen25technicalreport} (3B and 7B) deployed on local hardware, and DeepSeek \cite{deepseekai2024deepseekv3technicalreport,deepseekai2025deepseekr1incentivizingreasoningcapability} models (V3 and R1) accessed via API endpoints. Each graph is then evaluated for ontology conformance and KBQA performance.

\begin{table}[h]
    \centering
    \caption{Models for Graph Generation}
    \label{tab:model_deployment}
    \begin{tabular}{ll}
        \hline
        Model        & Deployment         \\
        \hline
        Qwen 2.5 3B  & Ollama on RTX 3060 \\
        Llama 3.2 3B & Ollama on RTX 3060 \\
        Qwen 2.5 7B  & LM Studio          \\
        Llama 3.1 8B & LM Studio          \\
        DeepSeek V3  & DeepSeek Platform  \\
        DeepSeek R1  & DeepSeek Platform  \\
        \hline
    \end{tabular}
\end{table}

KBQA evaluation was performed on Deepseek V3 via DeepSeek Platform.

@@

\begin{table}[h]
    \centering
    \caption{KBQA score (overall)}
    \label{tab:qa_percent}
    \begin{tabular}{|l|c|c|}
        \hline
        Graph       & Ontology-Free               & Ontology-Guided \\
        \hline
        Blank       & \multicolumn{2}{|c|}{60.50}                   \\
        \hline
        Qwen 2.5 3B & 73.50                       & 68.50           \\
        \hline
        Llama 3 3B  & 79.50                       & 78.00           \\
        \hline
        Qwen 2.5 7B & 95.50                       & 87.50           \\
        \hline
        Llama 3 8B  & 92.00                       & 90.50           \\
        \hline
        Deepseek V3 & \textbf{98.00}              & \textbf{97.50}  \\
        \hline
        Deepseek R1 & 97.50                       & 95.00           \\
        \hline
    \end{tabular}
\end{table}

\begin{table}[h]
    \centering
    \caption{KBQA score (single item)}
    \label{tab:qa_percent_single}
    \begin{tabular}{|l|c|c|}
        \hline
        Graph       & Ontology-Free               & Ontology-Guided \\
        \hline
        Blank       & \multicolumn{2}{|c|}{62.16}                   \\
        \hline
        Qwen 2.5 3B & 72.97                       & 67.57           \\
        \hline
        Llama 3 3B  & 80.54                       & 78.92           \\
        \hline
        Qwen 2.5 7B & 95.68                       & 89.19           \\
        \hline
        Llama 3 8B  & 94.05                       & 91.89           \\
        \hline
        Deepseek V3 & \textbf{100}                & \textbf{98.38}  \\
        \hline
        Deepseek R1 & 98.92                       & 96.22           \\
        \hline
    \end{tabular}
\end{table}

\begin{table}[h]
    \centering
    \caption{KBQA score (multi item)}
    \label{tab:qa_percent_multi}
    \begin{tabular}{|l|c|c|}
        \hline
        Graph       & Ontology-Free            & Ontology-Guided \\
        \hline
        Blank       & \multicolumn{2}{|c|}{40}                   \\
        \hline
        Qwen 2.5 3B & 80                       & 80              \\
        \hline
        Llama 3 3B  & 66.67                    & 66.67           \\
        \hline
        Qwen 2.5 7B & \textbf{93.33}           & 66.67           \\
        \hline
        Llama 3 8B  & 66.67                    & 73.33           \\
        \hline
        Deepseek V3 & 73.33                    & \textbf{86.67}  \\
        \hline
        Deepseek R1 & 80                       & 80              \\
        \hline
    \end{tabular}
\end{table}

@@

Comparing the ontology-free and guided approaches, while ontology-free graphs achieve marginally scores, the overall difference remains small. This close performance between ontology-free and ontology-guided graphs has important implications. It suggests that while enforcing ontological structure does not impede the extraction of answerable information, it also doesn't provide a substantial boost in question-answering accuracy in this specific evaluation setting. This could be due to several factors, including the nature of the questions asked, the effectiveness of the ontology integration in the prompting process, or the inherent ability of the LLMs to extract relevant information even without explicit ontological guidance.

@@

The practical value of the generated knowledge graphs was evaluated through a Knowledge Base Question Answering task. Interestingly, LLMs exhibited surprisingly strong baseline performance without the generated graphs, suggesting pre-existing knowledge of the museum's collections. While larger models showed improved KBQA accuracy, the impact of ontology guidance on this task was limited. This suggests that while crucial for semantic validity, ontology integration didn't significantly hinder factual information extraction in this evaluation. This also emphasizes the growing need for more sophisticated benchmarks that can better assess the benefits of ontology integration in KG generation from text.

In conclusion, our study contributes to the growing field exploring the synergy of LLMs, ontologies, and knowledge graphs within the cultural heritage domain. The high ontology conformance achieved by larger LLMs demonstrates the potential for creating structurally sound CIDOC-CRM knowledge graphs automatically. The KBQA results indicate their utility in augmenting information retrieval. Future research should focus on refining ontology integration techniques, exploring benefits in complex reasoning, and evaluating scalability across diverse archives. The overarching aim is to leverage these technologies to significantly enhance cultural heritage data management, accessibility, and research.

>>>>

\section{Introduction}
\label{sec:introduction}

@@

\section{Related Work}
\label{sec:related_work}

@@

\section{Methods}
\label{sec:methods}

@@

\section{Experiment}
\label{sec:experiment}

\begin{table}[h]
    \centering
    \caption{Models used for Knowledge Graph Generation} % Updated caption slightly
    \label{tab:model_deployment}
    \begin{tabular}{ll}
        \hline
        Model        & Deployment         \\
        \hline
        Qwen 2.5 3B  & Ollama on RTX 3060 \\
        Llama 3.2 3B & Ollama on RTX 3060 \\
        Qwen 2.5 7B  & LM Studio          \\
        Llama 3.1 8B & LM Studio          \\
        DeepSeek V3  & DeepSeek Platform  \\
        DeepSeek R1  & DeepSeek Platform  \\
        \hline
    \end{tabular}
\end{table}

We conducted experiments involving two main stages: \textbf{knowledge graph generation} and \textbf{knowledge graph evaluation}.

For the knowledge graph generation stage, we produced two sets of graphs from the collected textual dataset: one using an ontology-free approach and one using an ontology-guided approach. This generation was performed using open-weight LLMs of various sizes, specifically Llama 3 \cite{grattafiori2024llama3herdmodels} (3B and 8B) and Qwen 2.5 \cite{qwen2025qwen25technicalreport} (3B and 7B) deployed on local hardware, and DeepSeek \cite{deepseekai2024deepseekv3technicalreport,deepseekai2025deepseekr1incentivizingreasoningcapability} models (V3 and R1) accessed via API endpoints (Table \ref{tab:model_deployment}). In total, we generated twelve knowledge graphs: an ontology-free and an ontology-guided graph for each of the six listed models.

These constructed knowledge graphs were then subjected to evaluation. The intrinsic evaluation (Ontology Conformance), as described in Section \ref{sec:methods}, was performed on the ontology-guided graphs generated by each model. For the extrinsic evaluation, the Knowledge Base Question Answering (KBQA) task was conducted. In this task, the DeepSeek V3 model, accessed via the DeepSeek Platform, was employed as the question-answering model. For each question in the curated dataset, the DeepSeek V3 model was prompted to provide an answer based on one of the following contexts:
\begin{itemize}
    \item A verbalized representation of the ontology-free knowledge graph generated by one of the six distinct models
    \item A verbalized representation of the ontology-guided knowledge graph generated by one of the six distinct models
    \item A blank context, where the DeepSeek V3 model received only the question without any generated graph information
\end{itemize}

This setup allowed us to evaluate the question-answering performance when leveraging graphs generated by different models and approaches, comparing them against each other as well as a no-context "blank" scenario.

@@

\begin{table}[h]
    \centering
    \caption{KBQA score (overall)}
    \label{tab:qa_percent}
    \begin{tabular}{|l|c|c|}
        \hline
        Graph       & Ontology-Free               & Ontology-Guided \\
        \hline
        Blank       & \multicolumn{2}{|c|}{60.50}                   \\
        \hline
        Qwen 2.5 3B & 73.50                       & 68.50           \\
        \hline
        Llama 3 3B  & 79.50                       & 78.00           \\
        \hline
        Qwen 2.5 7B & 95.50                       & 87.50           \\
        \hline
        Llama 3 8B  & 92.00                       & 90.50           \\
        \hline
        DeepSeek V3 & \textbf{98.00}              & \textbf{97.50}  \\
        \hline
        DeepSeek R1 & 97.50                       & 95.00           \\
        \hline
    \end{tabular}
\end{table}

\begin{table}[h]
    \centering
    \caption{KBQA score (single item)}
    \label{tab:qa_percent_single}
    \begin{tabular}{|l|c|c|}
        \hline
        Graph       & Ontology-Free               & Ontology-Guided \\
        \hline
        Blank       & \multicolumn{2}{|c|}{62.16}                   \\
        \hline
        Qwen 2.5 3B & 72.97                       & 67.57           \\
        \hline
        Llama 3 3B  & 80.54                       & 78.92           \\
        \hline
        Qwen 2.5 7B & 95.68                       & 89.19           \\
        \hline
        Llama 3 8B  & 94.05                       & 91.89           \\
        \hline
        DeepSeek V3 & \textbf{100}                & \textbf{98.38}  \\
        \hline
        DeepSeek R1 & 98.92                       & 96.22           \\
        \hline
    \end{tabular}
\end{table}

\begin{table}[h]
    \centering
    \caption{KBQA score (multi item)}
    \label{tab:qa_percent_multi}
    \begin{tabular}{|l|c|c|}
        \hline
        Graph       & Ontology-Free            & Ontology-Guided \\
        \hline
        Blank       & \multicolumn{2}{|c|}{40}                   \\
        \hline
        Qwen 2.5 3B & 80                       & 80              \\
        \hline
        Llama 3 3B  & 66.67                    & 66.67           \\
        \hline
        Qwen 2.5 7B & \textbf{93.33}           & 66.67           \\
        \hline
        Llama 3 8B  & 66.67                    & 73.33           \\
        \hline
        DeepSeek V3 & 73.33                    & \textbf{86.67}  \\
        \hline
        DeepSeek R1 & 80                       & 80              \\
        \hline
    \end{tabular}
\end{table}

@@

Comparing the ontology-free and guided approaches, while ontology-free graphs achieve marginally higher scores, the overall difference remains small. This close performance between ontology-free and ontology-guided graphs has important implications. It suggests that while enforcing ontological structure does not impede the extraction of answerable information, it also doesn't provide a substantial boost in question-answering accuracy, at least in this specific evaluation setting. This could be due to several factors, including the nature of the questions asked, the effectiveness of the ontology integration in the prompting process, or the inherent ability of the LLMs to extract relevant information even without explicit ontological guidance.

@@

The practical value of the generated knowledge graphs was evaluated through a Knowledge Base Question Answering task. Interestingly, LLMs exhibited surprisingly strong baseline performance without the generated graphs, suggesting pre-existing knowledge of the museum's collections. While larger models showed improved KBQA accuracy, the impact of ontology guidance on this task was limited. This suggests that while crucial for semantic validity, ontology integration didn't significantly hinder factual information extraction in this evaluation. This also emphasizes the growing need for more sophisticated benchmarks that can better assess the quality of CIDOC-CRM knowledge graphs.

Our study contributes to the growing field exploring the synergy between LLMs, ontologies, and knowledge graphs within the cultural heritage domain. Our results demonstrate the potential for creating structurally sound and accurate CIDOC-CRM knowledge graphs automatically. However, we still notice much room for improvement in CIDOC-CRM knowledge graph evaluation methods, particularly in areas beyond the metrics measured in this study, such as entity duplication, a case where the same item appears twice in the same graph under different names. Future research could also explore other ontology integration techniques, refine datasets, and evaluate the scalability of these approaches across diverse archives. The overarching aim is to leverage these technologies to enhance cultural heritage data management, accessibility, and research.
